Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-05-21 manuscript/revision2/revision2.qmd
revision: Add Section S1 to the SI Appendix

We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the mash model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6). We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2). For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one. If rainfall amounts are uncorrelated, covariance will be near zero; if there is rank changing (e.g. high rainfall before flowering of A at MO but not MI, and high rainfall before flowering for B at MI but not MO), this covariance will be near negative one. We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model. The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B). If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the mash model.

@@

We used a greedy algorithm to select covariance matrices from each model's set that significantly improved the model likelihood. Five GxWeather covariance structures were selected, zero to two per subpopulation: two for green-up date, and three for flowering date. Of these five, three had mass on them in mash models of the strong effects (@fig-covar A-B). Two of these three matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while one had all positive or near-zero covariances. SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B). In total, 65% of the posterior weight of strong SNP effects in the mash model of Gulf green-up fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of green-up. The covariance matrix for this weather cue was visually similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A). Mash models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B). The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A). Only the Gulf subpopulation and Both subpopulations had mass on any GxWeather covariance matrices (@fig-covar C) for vegetative growth.

For flowering date, distinct GxWeather covariance structures captured covariance in effect sizes for SNPs in the Gulf and Midwest subpopulations. 33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (@fig-covar D). 22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (@fig-covar D), which had negative covariances between Texas and North gardens. Neither of these covariance matrices significantly improved the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had non-zero mass in models of Both subpopulations (@fig-covar E). In five of the six mash models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices. These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.

@@

We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B). We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in mash, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped one or more homologs from rice or *A. thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7). All flowering and green-up QTL intervals contained at least one SNP significant in at least one mash run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8). We also looked for enrichments of mash SNPs in the 1% tail of significance (the 'mash 1% tail') within each QTL interval. At the 5% level, three QTL had enrichments of SNPs in the mash 1% tail. Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the mash 1% tail in the QTL intervals. Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.

@@

```{=tex}
\clearpage
\onecolumn
\begin{table}[t!]
\centering
\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices. Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices \\
\midrule
1. cumulative GDD at 12C in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to green up date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to green up date & 60 \\
\bottomrule
\end{tabular}

\end{table}
\clearpage
\twocolumn
```
![Characterization of the timing of vegetative (green-up) and reproductive (flowering) growth in the switchgrass diversity panel. (a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens. Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation; map positions represent the original collection locations for the genotypes, and shapes represent the ecotype of the genotype. Vertical dashed lines indicate the summer solstice. Common gardens are arranged in latitudinal order. (b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations. (c) Narrow sense heritability of green-up and flowering within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations. (d) Flow diagram of the methods applied to the green-up and flowering dates to jointly estimate SNP effects across all sites. Mash was fit to SNP effect data and used to find covariance matrices that improved the mash model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a "strong" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png){#fig-map}

![Example hypothesis-driven covariance matrices specified in mash and the posterior weights placed on all covariance matrices. (a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1. Common gardens are arranged in latitudinal order within the matrices. Right column: Five example canonical covariance matrices. Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden. (b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date mash models, within and between two genetic subpopulations. Hypothesized covariance matrices (green). Covariance matrices included in mash that had zero posterior weight in all three mash runs on the genetic subpopulations, such as the identity matrix, are not shown. (c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png){#fig-covar}

@@

![Overlaps of QTL from an outbred pseudo-F2 cross and with jointly re-estimated SNP effects in the 1% tail of significance from a diversity panel. Dotted lines indicate permutation-based significance thresholds for each weather-related function. Stars indicate QTL with significant enrichment for SNPs in the 1% mash tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations. Rug plots show genomic locations of SNPs in the 1% mash tail for flowering date for each subpopulation. a) QTL mapping for the start of vegetative growth (Greenup Date), and three weather-related functions of greenup date. b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps.png){#fig-qtl}

@@

The formation and resequencing of the diversity panel has been described previously @lovell2021genomic. The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation. To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes [@kortefarlow2013], we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both'). Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic. Details on these models can be found in (SI Appendix, Section S3,S4).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel [@urbut2019flexible]. To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021]. We used the effect estimates for single nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect estimates for the two traits at each garden. We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure. We first modeled genetic effects with different combinations of covariance structures, using the effect size estimates from a subset of [19K]{.underline} relatively unlinked (r^2^ \< 0.2), randomly selected SNPs used in the GWAS. In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model. Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data. We then selected 19K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects), which gave sets of [**48K - 80K**]{.underline} SNPs, and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects. Notably, the majority of these "strong" effects were not significant in univariate GWAS.

We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (Table 1; SI Appendix, Table S1, Section S1). These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens. Mash SNP effects will undergo strong shrinkage towards one another in the first case, and little shrinkage in the second case. Mash also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data. We generated six data-driven matrices per mash run, five produced by singular value decomposition (SVD) of an overall matrix.

@@

To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals. The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S6. To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.

>>>>

We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6). We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2). For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one. If rainfall amounts are uncorrelated, covariance will be near zero; if there is rank changing (e.g. high rainfall before flowering of A at MO but not MI, and high rainfall before flowering for B at MI but not MO), this covariance will be near negative one. We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model. The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B). If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the *mash* model.

@@

We used a greedy algorithm to select covariance matrices from each model's set that significantly improved the model likelihood. Five GxWeather covariance structures were selected, zero to two per subpopulation: two for green-up date, and three for flowering date. Of these five, three had mass on them in *mash* models of the strong effects (@fig-covar A-B). Two of these three matrices had negative covariances between sites between Texas and North gardens (@fig-covar A), while one had all positive or near-zero covariances. SNP-associated phenotypic effects covaried with different weather-based cues in the Gulf & in Both subpopulations (@fig-covar B). In total, 65% of the posterior weight of strong SNP effects in the *mash* model of Gulf green-up fell on a covariance matrix constructed using the covariance of day length 14 days prior to the date of green-up. The covariance matrix for this weather cue was visually similar to the observed pattern of phenotypic correlation for the timing of vegetative growth in the Gulf subpopulation (@fig-map B; @fig-covar A). *Mash* models of the timing of vegetative growth in the Midwest subpopulation and Both subpopulations did not include this GxWeather covariance; the Midwest had no weight on any GxWeather covariance, while Both subpopulations had non-zero weights on two additional GxWeather covariance types, average temperature one day prior to green-up, and the day length change in seconds in the day prior to green-up (@fig-covar B). The average temperature covariance matrix had negative covariances between Texas and North gardens, though not as strong as the negative phenotypic correlations seen in Both subpopulations (@fig-map B; @fig-covar A). Only the Gulf subpopulation and Both subpopulations had mass on any GxWeather covariance matrices (@fig-covar C) for vegetative growth.

For flowering date, distinct GxWeather covariance structures captured covariance in effect sizes for SNPs in the Gulf and Midwest subpopulations. 33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (@fig-covar D). 22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (@fig-covar D), which had negative covariances between Texas and North gardens. Neither of these covariance matrices significantly improved the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had non-zero mass in models of Both subpopulations (@fig-covar E). In five of the six *mash* models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices. These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.

@@

We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B). We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped one or more homologs from rice or *A. thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7). All flowering and green-up QTL intervals contained at least one SNP significant in at least one *mash* run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8). We also looked for enrichments of *mash* SNPs in the 1% tail of significance (the '*mash* 1% tail') within each QTL interval. At the 5% level, three QTL had enrichments of SNPs in the *mash* 1% tail. Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the *mash* 1% tail in the QTL intervals. Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.

@@

```{=tex}
\clearpage
\onecolumn
\begin{table}[t!]
\centering
\caption{Weather variables and time frames prior to the start of vegetative and reproductive growth used to construct the hypothesis-based covariance matrices. Covariance matrices were constructed for, and tested on, three subpopulations and two phenological traits. The correlations between values of these weather variables for genetically identical plants grown in different gardens were used to fill off-diagonal cells of the covariance matrices. Narrow-sense heritabilities for these values at each garden were used for the diagonal cells.}
\begin{tabular}{lrr}
Weather variable & Time Frame & Covariance Matrices \\
\midrule
1. cumulative growing degree days at 12C in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
2. cumulative rainfall in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60  \\
3. day length (hours) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
4. day length change (seconds) on a specific day indicated by the time frame & (1-7) and 14 days prior to the phenological date & 48 \\
5. average temperature in the time frame & (1-7), 14, 21, 28 days prior to the phenological date & 60 \\
\bottomrule
\end{tabular}

\end{table}
\clearpage
\twocolumn
```
![Characterization of the timing of vegetative (green-up) and reproductive (flowering) growth in the switchgrass diversity panel. (a) Map and trait histograms of green-up and flowering dates across two genetically distinct switchgrass subpopulations and eight common gardens. Purple represents individuals from the Midwest genetic subpopulation, and pink individuals from the Gulf subpopulation; map positions represent the original collection locations for the genotypes, and shapes represent the ecotype of the genotype. Vertical dashed lines indicate the summer solstice. Common gardens are arranged in latitudinal order. (b) Phenotypic correlations between clonal replicates planted at eight common gardens, within and between two genetic subpopulations. (c) Narrow sense heritability of green-up and flowering within single common gardens (purple) and across all eight common gardens (green), within and between two genetic subpopulations. (d) Flow diagram of the methods applied to the green-up and flowering dates to jointly estimate SNP effects across all sites. *Mash* was fit to SNP effect data and used to find covariance matrices that improved the *mash* model likelihood using a large set of randomly selected, relatively unlinked SNP effects; this model was applied to a "strong" set of SNP effects with large effect sizes in the univariate GWAS.](images/Fig1_200dpi.png){#fig-map}

![Example hypothesis-driven covariance matrices specified in *mash* and the posterior weights placed on all covariance matrices. (a) Left column: Five example hypothesized covariance matrices specified for the green-up date or flowering date phenotype; these matrices were created from environment-specific correlations across eight common gardens, and are described in Table 1. Common gardens are arranged in latitudinal order within the matrices. Right column: Five example canonical covariance matrices. Canonical matrices (purple) have simple interpretations, such as equal effects across all common gardens, or effects specific to a single common garden. (b,d) Total posterior weight placed on each covariance matrix type specified for (b) green-up date and (d) flowering date *mash* models, within and between two genetic subpopulations. Hypothesized covariance matrices (green). Covariance matrices included in *mash* that had zero posterior weight in all three *mash* runs on the genetic subpopulations, such as the identity matrix, are not shown. (c,e) Total posterior weight placed on covariance matrices that were hypothesized or canonical, for the (c) green-up date phenotype and (e) flowering date phenotype.](images/Figure_2_recolor_covariance_matrices_and_posterior_weights.png){#fig-covar}

@@

![Overlaps of QTL from an outbred pseudo-F2 cross and with jointly re-estimated SNP effects in the 1% tail of significance from a diversity panel. Dotted lines indicate permutation-based significance thresholds for each weather-related function. Stars indicate QTL with significant enrichment for SNPs in the 1% *mash* tail; G, M, and B indicate which subpopulation had enrichment: G - Gulf subpopulation, M -Midwest subpopulation, B - both subpopulations. Rug plots show genomic locations of SNPs in the 1% *mash* tail for flowering date for each subpopulation. a) QTL mapping for the start of vegetative growth (Greenup Date), and three weather-related functions of greenup date. b) QTL mapping for the start of reproductive growth (Flowering Date), and two weather-related functions of flowering date.](images/Figure_4_QTL_Overlaps.png){#fig-qtl}

@@

The formation and resequencing of the diversity panel has been described previously @lovell2021genomic. The diversity panel contained 134 sequenced, clonally propagated individuals from the Midwest genetic subpopulation, and 229 from the Gulf genetic subpopulation. To allow for the possibility that different subpopulations had different strengths of connection between our phenotypes and genotypes [@kortefarlow2013], we conducted three sets of genetic analyses: on Gulf and Midwest genotypes separately, and on both subpopulations together ('Both' subpopulations). Analyses to determine narrow-sense heritability (h^2^) for green-up and flowering were done using linear mixed models and followed @lovell2021genomic. Details on these models can be found in (SI Appendix, Section S2).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP effects across our common gardens, we used multivariate adaptive shrinkage (*mash*) on SNP effect estimates from the diversity panel [@urbut2019flexible]. To do this, we first conducted univarate genome-wide association (GWAS) on site-specific best linear unbiased predictors (BLUPs) for the timing of vegetative growth and flowering, using the switchgrassGWAS package and the methods in [@lovell2021]. We used the effect estimates for single nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect estimates for the two traits at each garden. We then jointly modeled genetic effect estimates, and the mixture of ways these effects might covary, across all eight common gardens, using a two-part procedure. We first modeled genetic effects with different combinations of covariance structures, using the effect size estimates from a subset of [19K]{.underline} relatively unlinked (r^2^ \< 0.2), randomly selected SNPs used in the GWAS. In these models (hereafter referred to as "random"), we then specified a variety of potential covariance structures among genetic effect size estimates from each garden (as expanded on below) and used a greedy algorithm to iteratively select and add covariance structures that significantly improved the log likelihood of the *mash* model. Details on these models can be found in (SI Appendix, Section S3). Given the high correlation between some of the covariance patterns we tested, this approach allowed us to select only the subset of patterns of covariance that significantly improved the fit of the jointly inferred structure of genetic effect estimates present in the data. We then selected 19K relatively unlinked (r^2^ \< 0.2) SNPs from each univariate GWAS that had the largest effect estimates in each GWAS (hereafter "strong" effects), which gave sets of [**48K - 80K**]{.underline} SNPs, and jointly re-estimated these effect size estimates at all eight sites using the set of significant covariance matrices that most improved the model fit on the random effects. Notably, the majority of these "strong" effects were not significant in univariate GWAS.

We generated hypothesis-based covariance matrices derived from correlations in environmental cues in the green-up or flowering date windows for the three populations (Table 1; SI Appendix, Table S1, Section S1). These covariance matrices represent correlations between identical genotypes drawn from a specific population at pairs of common gardens; covariances near one mean that the population has a strong, positive linear relationship in individual responses at that pair of gardens, while covariances near zero mean that there is no relationship within the population for individual responses at that pair of gardens. *Mash* SNP effects will undergo strong shrinkage towards one another in the first case, and little shrinkage in the second case. *Mash* also generates data-driven covariance matrices corresponding to major patterns of SNP effects present in the data. We generated six data-driven matrices per *mash* run, five produced by singular value decomposition (SVD) of an overall matrix.

@@

To confirm candidate genomic regions and patterns of allelic effects found in the diversity panel, we analyzed flowering in an outbred pseudo-F2 cross between four individuals, two Midwest and two Gulf individuals. The formation of this mapping population has been described previously [@milano2016genetic]; additional details on QTL mapping can be found in SI Appendix, Section S4. To be directly comparable to the diversity panel data, only 2019 phenology data from the pseudo-F2 cross from the same eight common garden sites were used.
